Probability of causal contact between interstellar civilizations through MonteCarlo simulations

### 2019-08-21 manuscript.tex
detalles

Since the method is based on events, the first step is to define an
architecture of events, the relationships between them and how events
trigger changes on the state of the system.
%
We consider a simulated system that represents the galactic habitable
zone (GHZ).
%
On a first approach, the GHZ is a 2-dimensional annular region.
%
The adopted values for the GHZ are XX and XX \citep{Lineweaver2004}.
%
This simple model does not take into accont the variations in stellar
density given by the spiral structure.
%
Although there are several possible approaches, we chose to follow the
evolution of the system according to the following events:

@@

We implemented the simulation of a regular grid of models varying over the
hypothesis space, which covers XXX different models.
%
For each model, we simulated XXX realizations with different random
seeds, in order to estimate the variance.
%
The simulated parameters, $\lambda_1$, $\lambda_2$ and $D_{max}$ cover
the range XX-XX, XX-XX and XX-XX, respectively, with a partition of XX
values for each parameter.

As a product of the simulations, several quantities are obtained.
%
Some quantities are directly derived from the discrete events, namely,
the ID of emitting CETI, the ID of emitting (receiving) CETI, the
position in the galaxy, the time of Awakening (Contact) and the time
of Doomsday  (Blackout) for each CETI.
%
This structure is always for the A-D cycle of each emitting CETI, and
repeats for each contact (C-B).
%
We can also derive the additional quantities.
%
The age of a CETI is the time elapsed from A to D.
%
%
%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
%the CETI receiving signals (receiver), ceti\_c is the ceti that
%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
%that is lestened by at least another CETI.
%
The time span of a CETI listening another or being listened by another
can also be easily derived from the results of a realization.
%
This way we can also compute the distribution in the galaxy of cetis
that reach contact and the waiting time until the first contact or the
waiting time until the next contact.
%
Regarding the properties of the population of CETIs and its evolution,
we analyze the fraction of awaken time a ceti is listening at least
another ceti, the age of contacted CETIs at first contact, the
distribution of time to wait until next contact, the fraction of cetis
where the first contact is given at the awakening, the distribution of
the number of contacts for each CETI, the distribution of the maximum
number of contacts for each CETI, the distribution of the number of
contacts as a function of CETI age, the number of contacts as a
function of time in the galaxy, the rate of CETIs that succeed in
contact, and the distribution of distances between contacted CETIs.
%
Usefull derived quantity is the duration of two-way communication
channels the or the fraction of contacts that admit a response.
% 
We also analyze the relations between the distance to ceti vs. time of
double communication, the distance to ceti vs. age of contacted ceti,
the age of a ceti and the maximum number of contacted cetis before D,
or the lifespan of a ceti vs. max number of contacts.
%
All these quantities can be analyzed as a function of the simulation
parameters.
%

We chose eight models that are on fairly oposed regions of the
explored parameter space, and cover short/long lifetimes, short/long
waiting times for the next CETI to appear in the Galaxy and
short/large range reach of the signal, including all possible
combinations.
%
The details of each model are summarized in the
Table~\ref{T_selected_models}.

@@

In what follows, we focus on the analysis of the duration of contacts
(Sec. X), the waiting time for the first contact (Sec. X) the
variations of contact posibilities as a function of the position of
the CETI in the Galaxy (Sec. X) and the occurrence of reciprocal
communication (Sec. X).

@@

On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
zero, one, two or three contacts, as a function of the galactocentric
distance.
%
For this model, which corresponds to XXX, the majority of CETIs do not
make any contact during all the time period in which they are active.
%
As expected, the fractions of CETIs making at least one contact
diminishes as the number of contacts rises.
%
The shape of this distribution is a result of the geometry of the GHZ.
%
In general, a closer position to the center of the Galaxy slightly favours 
the posibility of a contact.
%
This would indicate that the position of the Earth is suitable for the
odds of making a contact.

>>>>

Since the method is based on events, the first step is to define an
architecture of events, the relationships between them and how events
trigger changes on the state of the system.
%
We consider a simulated system that represents the galactic habitable
zone (GHZ).
%
On a first approach, the GHZ is a 2-dimensional annular region.
%
The adopted values for the GHZ are 20 kly and 60 kly, for the inner
and outer radius, respectively \citep{Lineweaver2004}.
% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
% mas chico...
%
This simple model does not take into accont the variations in stellar
density given by the spiral structure.
%
Although there are several possible approaches, we chose to follow the
evolution of the system according to the following events:

@@

We implemented the simulation of a regular grid of models varying over
the hypothesis space, which covers 15000 different models.
%
For each model, we simulated 50 realizations with different random
seeds, in order to estimate the variance.
%
The simulated parameters, $\tau_a$, $\tau_s$ cover the ranges
4000-200000 and 10000-500000, respectively, with a partition of 50
values for each parameter.
%
For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
20000, 40000 and 80000 lyr.
%
This setup makes a total of 750000 simulations, each covering a time
range of one million years.

As a product of the simulations, several quantities are obtained.
%
Some quantities are directly derived from the discrete events, namely,
the ID of emitting CETI, the ID of emitting (receiving) CETI, the
position in the galaxy, the time of Awakening (Contact) and the time
of Doomsday  (Blackout) for each CETI.
%
This structure is always for the A-D cycle of each emitting CETI, and
repeats for each contact (C-B).
%
We can also derive the additional quantities.
%
The age of a CETI is the time elapsed from A to D.
%
%
%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
%the CETI receiving signals (receiver), ceti\_c is the ceti that
%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
%that is lestened by at least another CETI.
%
The time span of a CETI listening another or being listened by another
can also be easily derived from the results of a realization.
%
This way we can also compute the distribution in the galaxy of cetis
that reach contact and the waiting time until the first contact or the
waiting time until the next contact.
%
Regarding the properties of the population of CETIs and its evolution,
we analyze the fraction of awaken time a ceti is listening at least
another ceti, the age of contacted CETIs at first contact, the
distribution of time to wait until next contact, the fraction of cetis
where the first contact is given at the awakening, the distribution of
the number of contacts for each CETI, the distribution of the maximum
number of contacts for each CETI, the distribution of the number of
contacts as a function of CETI age, the number of contacts as a
function of time in the galaxy, the rate of CETIs that succeed in
contact, and the distribution of distances between contacted CETIs.
%
Usefull derived quantity is the duration of two-way communication
channels the or the fraction of contacts that admit a response.
% 
We also analyze the relations between the distance to ceti vs. time of
double communication, the distance to ceti vs. age of contacted ceti,
the age of a ceti and the maximum number of contacted cetis before D,
or the lifespan of a ceti vs. max number of contacts.
%
All these quantities can be analyzed as a function of the simulation
parameters.
%
We chose eight models that are on fairly oposed regions of the
explored parameter space, and cover short/long lifetimes, short/long
waiting times for the next CETI to appear in the Galaxy and
short/large range reach of the signal, including all possible
combinations.
%
The details of each model are summarized in the
Table~\ref{T_selected_models}.

@@

In what follows, we focus on the analysis of the duration of contacts
(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
(Sec.~\ref{SS_waiting}) and the
variations of contact posibilities as a function of the position of
the CETI in the Galaxy (Sec.~\ref{SS_location}).

@@

On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
zero, one, two or three contacts, as a function of the galactocentric
distance, for model M1.
% VERIFICAR!  M1?
%
For this model, the majority of CETIs do not
make any contact during all the time period in which they are active.
%
As expected, the fractions of CETIs making at least one contact
diminishes as the number of contacts rises.
%
The shape of this distribution is a result of the geometry of the GHZ.
%
In general, a closer position to the center of the Galaxy slightly favours 
the posibility of a contact.
%
This would indicate that the position of the Earth is suitable for the
odds of making a contact.
